Simple Macroeconomic Forecast Distributions

### 2024-03-18 presentation/presentation_svr.tex
presentation files

\title[Simple Prediction Intervals]{Simple Macroeconomic Forecast Distributions}
\subtitle{} 
\author[Friederike Becker, Fabian Krüger, Melanie Schienle]{Friederike Becker, Fabian Krüger, Melanie Schienle}

\date[05.\,12.\,2023]{December 5, 2023}

@@

\section{Setting}

@@

@@

\begin{columns}
\begin{column}{0.4\textwidth}
   	\begin{itemize}
         \item usual practice: issue point forecasts only
         \begin{itemize}
	    \item uncertainty is at best acknowledged, rarely quantified \bigskip
	\end{itemize}
        %\item media reception often fixated on zero
        \item forecasts of different horizons are often left uncontextualized \bigskip
        \item distributional forecasts supposedly would require extra modeling effort
    \end{itemize}
    \vspace{2cm}
\end{column}
\begin{column}{0.6\textwidth}
    \begin{figure}
        \centering
        \includegraphics[width=0.8\textwidth]{figures/recession_light_underlined_green_smaller_date.jpg} 
        \caption{Source: tagesschau.de}
        \label{fig:enter-label}
    \end{figure}     
\end{column}
\end{columns}
\end{frame}

\begin{frame}[t]{Contributions of this work}
\begin{itemize}
    \item show that attaching prediction intervals to an existing base of point forecasts can be 
    \begin{itemize} 
    \item simple
    \item cheap
    \item transparent
    \end{itemize}   
    \item provide competitively performing distributional forecasts for 
    \begin{itemize}
    \item  GDP growth and inflation
    \item  all G7 countries
    \item  current and next year targets
    \end{itemize}
    %\item simple and transparent method of attaching prediction intervals 
%\item show that attaching distributional forecasts to an existing
%\item for up to one and a half years into the future 
%\item evaluate on years 1999-2021
\end{itemize}
\end{frame}

@@

\section{Methods}
\begin{frame}{Methods}
We apply an attractively simple and cheap method. For a given country and target:
\begin{itemize}
    \item given forecasts $\hat{y}_{t, h}$ and the realized true values $y_{t}$ ...
    \begin{itemize}
        \item for target year $t$, horizon $h$
    \end{itemize}
    \item ... construct sets $\mathcal{E}_{t, h} = \{\hat{e}_{t^*, h}  | t-R \leq t^* < t \}$, containing the last $R$ forecast errors
    \begin{itemize}
        \item based on \textit{absolute} errors $\hat{e}_{t,h} = |y_{t} - \hat{y}_{t, h}| $
        \item currently: $R=9$
        %\item sets are stratified by country, target and horizon
    \end{itemize}
    \item for $\alpha \in \{0.5, 0.8\}$, compute $q^{\alpha}_{t, h} =  Q\left(\mathcal{E}_{t, h}, \alpha \right)$
    \item and compute the upper and lower endpoints of a central prediction interval as 
    \begin{itemize}
        \item $u^{\alpha}_{t,h} = \hat{y}_{t, h} + q^{\alpha}_{t, h}$ 
        \item $l^{\alpha}_{t,h} \hspace{1mm} = \hat{y}_{t, h} - q^{\alpha}_{t, h}$
    \end{itemize}
    \item use PAVA-type reordering if intervals shrink with increasing horizon
    \item assess central interval coverage, score via the weighted interval score (WIS)\footnote{Bracher, J. et al. 2021. \textit{Evaluating
Epidemic Forecasts in an Interval Format.} PLoS Computational Biology 17 (2)}
\end{itemize}

@@

\section{Results}
\begin{frame}{Scores }
%\textit{I'll make the figures, incl. legends, nicer and also remove (the reason for) the joke about the devoured overprediction}
    \begin{figure}
        \centering
        \includegraphics[width=0.95\textwidth]{figures/wis_cpigdp.pdf} 
        \label{fig:enter-label}
    \end{figure} 
\end{frame}

\begin{frame}{Calibration - Interval Coverage Levels}
%\textit{Here, I'll instead map the colors to the different forecast sources (currently mapped to the window method), so we'll just have one plot each}
    %Coverage at the 50\% and 80\% central interval: \\ 
    \begin{figure}
        \centering
        \includegraphics[width=0.9\textwidth]{figures/coverage.pdf}
        %\caption{Coverage at the 50 and 80 percent level}
        \label{fig:enter-label}
    \end{figure}
\end{frame}

\begin{frame}{Increasing Uncertainty}
%Width of Prediction Intervals, by Horizon
%\textit{include graphic or table showing increasing uncertainty with rising horizons}\\ 
\begin{columns}
\begin{column}{0.6\textwidth}
    \begin{figure}
        \centering
        \includegraphics[width=\textwidth]{figures/horizon_uncc.pdf}
        %\caption{Coverage at the 50 and 80 percent level}
        \label{fig:enter-label}
    \end{figure}
\end{column}
\begin{column}{0.4\textwidth}
\centering
\textbf{Average length of intervals}
\begin{table}
\begin{tabular}{ l c c }
&   50\%  & 80\%\\[0.3em]
\multicolumn{3}{c}{\textbf{GDP Growth}}\\
Fall,  SY & 0.5 & 1.1\\ 
Spring, SY &1.2 &2.1 \\ 
Fall, NY & 1.9 & 3.5\\ 
Spring, NY &2.2 &4.0 \\[0.3em] 
\multicolumn{3}{c}{\textbf{Inflation}}\\
%&   50\%  & 80\%\\[0.3em]
Fall,  SY & 0.3 & 0.5\\ 
Spring, SY &0.6 &1.1 \\ 
Fall, NY & 1.1 & 1.8\\ 
Spring, NY &1.3 &2.1 \\ 
\end{tabular}
\end{table}
\vspace{1.25cm}
\end{column}
\end{columns}

@@

\appendix
\beginbackup
\begin{frame}{Directional Errors - Scores}
\label{errorextraction}
\begin{figure}
        \centering
        \includegraphics[width=0.9\textwidth]{figures/wis_cpigdp_directional.pdf}
       %\caption{Coverage at the 50 and 80 percent level}
    \end{figure}
\end{frame}
\begin{frame}{Directional Errors - Coverage}
\begin{figure}
        \centering
        \includegraphics[width=0.9\textwidth]{figures/coverage_directional.pdf}
        %\caption{Coverage at the 50 and 80 percent level}
    \end{figure}
\end{frame}

@@

\end{document}

>>>>

\title[Simple Prediction Intervals]{Simple Macroeconomic Forecast Distributions}
%\subtitle{Einfache makroökonomische probabilistische Vorhersagen} 
\author[Friederike Becker, Fabian Krüger, Melanie Schienle]{Friederike Becker, Fabian Krüger, Melanie Schienle}

\date[28.\,02.\,2024]{February 28, 2024}

@@

\begin{frame}{Small Note on Terminology}
Some terms are used interchangeably within the context of this presentation
\begin{itemize}
\item probabilistic - distributional
\item forecast - prediction - projection 
\end{itemize}
\end{frame}

\section{Setting and Motivation}

@@

\begin{frame}[t]{Why should we explicitly quantify uncertainty?}
\begin{itemize}
    \item more honest and transparent \smallskip
    \item can increase trust in forecasts \smallskip
    \item broader information base facilitates better decision making \smallskip
    \item better evaluation and accountability
    %\item simple and transparent method of attaching prediction intervals 
%\item show that attaching distributional forecasts to an existing
%\item for up to one and a half years into the future 
%\item evaluate on years 1999-2021
\end{itemize}
\end{frame}

@@

\begin{columns}
\begin{column}{0.45\textwidth}
   	\begin{itemize}
         \item usual practice: issue point forecasts only
         \begin{itemize}
	    \item uncertainty is at best acknowledged, rarely quantified \smallskip
	\end{itemize}
        %\item media reception often fixated on zero
        \item we observe artefacts in reporting 
        \begin{itemize}
            \item fixation on values slightly above/below zero
            \item different horizons are often left uncontextualized \bigskip
        \end{itemize}
        \item distributional forecasts supposedly require additional ...
        \begin{itemize}
        \item ... modeling effort (forecast issuer)
        \item ... cognitive load (forecast user)
        \end{itemize}
    \end{itemize}
    \vspace{2cm}
\end{column}
\begin{column}{0.55\textwidth}
\vspace*{-2.05cm}
    \begin{figure}
        \centering
        \includegraphics[width=0.8\textwidth]{figures/recession_light_underlined_green_smaller_date.jpg} 
        \caption{Source: tagesschau.de}
        \label{fig:enter-label}
    \end{figure}     
\end{column}
\end{columns}
\end{frame}

\begin{frame}[t]{Types of probabilistic forecasts}

\begin{columns}
\begin{column}{0.39\textwidth}
    \begin{itemize}
\item full probability distribution
\item histogram
\item samples from distribution
\item \textbf{quantiles or prediction intervals}
\end{itemize}
\begin{figure}
        \centering
        \includegraphics[width=0.8\textwidth]{figures_svr/densityfc.pdf} 
        \label{fig:enter-label}
    \end{figure}
\end{column}
\begin{column}{0.52\textwidth}
\vspace*{-1.55cm}
    \begin{figure}
        \centering
        \includegraphics[width=0.7\textwidth]{figures_svr/CI_illustration.pdf} 
        \label{fig:enter-label}
    \end{figure}     
\end{column}
\begin{column}{0.01\textwidth}

\end{column}
\end{columns}
\end{frame}

\addtocounter{framenumber}{-1}
\begin{frame}[t]{Example 1: Weather Forecasting}
\begin{columns}
\begin{column}{0.35\textwidth}
Uncertainty is communicated via \smallskip
    \begin{itemize}
         \item $80\%$ prediction interval for temperature \medskip
         \item probability for precipitation and adverse weather events
    \end{itemize}
    \vspace{2cm}
\end{column}
\begin{column}{0.65\textwidth}
\vspace*{-.05cm}
    \begin{figure}
        \centering
        \includegraphics[width=0.8\textwidth]{figures_svr/Probcast.jpg} 
        \caption{Source: Mass, C. et al. 2009. \textit{PROBCAST}, BAMS}
    \end{figure}
\end{column}
\end{columns}
\end{frame}

\addtocounter{framenumber}{-1}
\begin{frame}[t]{Example 2: Forecasting Covid-19}
\begin{columns}
\begin{column}{0.35\textwidth}
Uncertainty is communicated via \smallskip
    \begin{itemize}
         \item granular prediction intervals for Covid-19 Cases and Deaths \medskip
         \item main prediction intervals ($50\%$, $95\%$) are highlighted
    \end{itemize}
    \vspace{2cm}
\end{column}
\begin{column}{0.65\textwidth}
\vspace*{-.05cm}
    \begin{figure}
        \centering
        \includegraphics[width=0.65\textwidth]{figures_svr/covid19.jpg} 
        \caption{Source: https://viz.covid19forecasthub.org}
    \end{figure}
\end{column}
\end{columns}
\end{frame}

\begin{frame}[t]{The makings of a good probabilistic forecast}
\vspace*{-0.25cm}
\begin{itemize}
\item central paradigm: \textbf{sharpness subject to calibration}\footnote{Gneiting, T. et al. 2007. \textit{Probabilistic forecasts, calibration and sharpness.} JRSS}
\item calibration alone can be easily asserted
\begin{itemize}
\item most important metric for a trustworthy forecast
\end{itemize}
\item \textbf{proper scoring rules} check calibration and sharpness simultaneously
\end{itemize}
\begin{figure}
        \centering
        \includegraphics[width=0.61\textwidth]{figures_svr/sharp_calib.pdf} 
    \end{figure}
\end{frame}

\section{Methods}
\begin{frame}[t]{Contributions of this work}
\begin{itemize}
    \item show that attaching prediction intervals to an existing base of point forecasts can be 
    \begin{itemize} 
    \item simple
    \item cheap
    \item transparent
    \end{itemize}   
    \item provide competitively performing distributional forecasts for 
    \begin{itemize}
    \item  GDP growth and inflation
    \item  all G-7 countries
    \item  current and next year targets
    \end{itemize}
    %\item simple and transparent method of attaching prediction intervals 
%\item show that attaching distributional forecasts to an existing
%\item for up to one and a half years into the future 
%\item evaluate on years 1999-2021
\end{itemize}
\end{frame}

@@

%\section{Methods}
\begin{frame}{Methods}
We apply an attractively simple and cheap method. For a given country and target:
\begin{itemize}
    \item given forecasts $\hat{y}_{t, h}$ and the realized true values $y_{t}$ ...
    \begin{itemize}
        \item for target year $t$, horizon $h$
    \end{itemize}
    \item ... construct sets $\mathcal{E}_{t, h} = \{\hat{e}^{a}_{t^*, h}  | t-R \leq t^* < t \}$, containing the last $R$ forecast errors
    \begin{itemize}
        \item based on \textit{absolute} errors $\hat{e}^{a}_{t,h} = |y_{t} - \hat{y}_{t, h}| $
        \item currently: $R=11$
        %\item sets are stratified by country, target and horizon
    \end{itemize}
    \item for each desired confidence level $\tau$, compute $q^{\tau}_{t, h} =  Q\left(\mathcal{E}_{t, h}, \tau \right)$
    \begin{itemize}
    \item  for conciseness, we choose $\tau  \in \{0.5, 0.8\}$
    \end{itemize}
    \item and compute the upper and lower endpoints of a central prediction interval as 
    \begin{itemize}
        \item $u^{\tau}_{t,h} = \hat{y}_{t, h} + q^{\tau}_{t, h}$ 
        \item $l^{\tau}_{t,h} \hspace{1mm} = \hat{y}_{t, h} - q^{\tau}_{t, h}$
    \end{itemize}
    %\item use PAVA-type reordering if intervals shrink with increasing horizon

\begin{frame}{Methods, continued}
\begin{itemize}
\item Notwithstanding the method's simplicity, the following constraints are imposed
\begin{itemize}
    \item point forecast lies at the center of each prediction interval  
    \item intervals are reordered if their length does not increase with distance to target 
\end{itemize}
\item as an alternative, ``directional'' errors are also considered \bigskip
\item assess central interval coverage and score via the weighted interval score (WIS)\footnote{Bracher, J. et al. 2021. \textit{Evaluating
Epidemic Forecasts in an Interval Format.} PLoS Computational Biology 17 (2)}
\begin{equation*} \label{eq:interval_score}
IS_{\tau}(F, y) = (u-l) + \frac{2}{1-\tau}(l - y)1(y < l) + \frac{2}{1-\tau}(y - u)1(y > u).
\end{equation*}
\end{itemize}
\end{frame}

@@

\section{Results}
\begin{frame}{Scores - Inflation}
%\textit{I'll make the figures, incl. legends, nicer and also remove (the reason for) the joke about the devoured overprediction}

\begin{columns}
\begin{column}{0.35\textwidth}
    \begin{itemize}
         \item IMF-based forecasts perform relatively well 
         \begin{itemize}
         \item more so at long horizons\medskip
         \end{itemize}
         \item overprediction contributes most to scores
    \end{itemize}
    \vspace{2cm}
\end{column}
\begin{column}{0.65\textwidth}
\vspace*{-.95cm}
    \begin{figure}
        \centering
        \includegraphics[width=0.8\textwidth]{figures_svr/svr_wis_pcpi_pch_absolute_rollingwindow.pdf} 
        \label{fig:enter-label}
    \end{figure} 
\end{column}
\end{columns}
\end{frame}

\begin{frame}{Scores - GDP Growth}
%\textit{I'll make the figures, incl. legends, nicer and also remove (the reason for) the joke about the devoured overprediction}
\begin{columns}
\begin{column}{0.35\textwidth}
    \begin{itemize}
         \item IMF-based forecasts perform slightly better 
         \begin{itemize}
         \item across all horizons\medskip
         \end{itemize}
         \item dispersion and overprediction contribute most to scores
    \end{itemize}
    \vspace{2cm}
\end{column}
\begin{column}{0.65\textwidth}
\vspace*{-.95cm}
    \begin{figure}
        \centering
        \includegraphics[width=0.8\textwidth]{figures_svr/svr_wis_ngdp_rpch_absolute_rollingwindow.pdf} 
        \label{fig:enter-label}
    \end{figure} 
\end{column}
\end{columns}
\end{frame}

%\begin{frame}{Calibration - Interval Coverage Levels}
%\textit{Here, I'll instead map the colors to the different forecast sources (currently mapped to the window method), so we'll just have one plot each}
    %Coverage at the 50\% and 80\% central interval: \\

%    \begin{figure}
%        \centering
%        \includegraphics[width=0.9\textwidth]{figures/coverage.pdf}
%        %\caption{Coverage at the 50 and 80 percent level}
%        \label{fig:enter-label}
%    \end{figure}
%\end{frame}

\begin{frame}{Calibration - Interval Coverage Levels - Inflation}
%\textit{Here, I'll instead map the colors to the different forecast sources (currently mapped to the window method), so we'll just have one plot each}
    %Coverage at the 50\% and 80\% central interval: \\ 
    \begin{figure}
        \centering
        \includegraphics[width=0.98\textwidth]{figures_svr/svr_coverage_pcpi_pch_absolute_rollingwindow.pdf}
        %\caption{Coverage at the 50 and 80 percent level}
        \label{fig:enter-label}
    \end{figure}
\end{frame}

\begin{frame}{Calibration - Interval Coverage Levels - GDP Growth}
%\textit{Here, I'll instead map the colors to the different forecast sources (currently mapped to the window method), so we'll just have one plot each}
    %Coverage at the 50\% and 80\% central interval: \\ 
    \begin{figure}
        \centering
        \includegraphics[width=0.98\textwidth]{figures_svr/svr_coverage_ngdp_rpch_absolute_rollingwindow.pdf}
        %\caption{Coverage at the 50 and 80 percent level}
        \label{fig:enter-label}
    \end{figure}
\end{frame}

\begin{frame}{Increasing Uncertainty}
%Width of Prediction Intervals, by Horizon
%\textit{include graphic or table showing increasing uncertainty with rising horizons}\\ 
\begin{columns}
\begin{column}{0.6\textwidth}
    \begin{figure}
        \centering
        \includegraphics[width=\textwidth]{figures/horizon_uncc_new.pdf}
        %\caption{Coverage at the 50 and 80 percent level}
        \label{fig:enter-label}
    \end{figure}
\end{column}
\begin{column}{0.4\textwidth}
\centering
\textbf{Average length of intervals}
\begin{table}
\begin{tabular}{ l c c }
&   50\%  & 80\%\\[0.3em]
\multicolumn{3}{c}{\textbf{GDP Growth}}\\
Fall,  SY & 0.5 & 1.1\\ 
Spring, SY &1.2 &2.1 \\ 
Fall, NY & 1.9 & 3.5\\ 
Spring, NY &2.2 &4.0 \\[0.3em] 
\multicolumn{3}{c}{\textbf{Inflation}}\\
%&   50\%  & 80\%\\[0.3em]
Fall,  SY & 0.3 & 0.5\\ 
Spring, SY &0.6 &1.1 \\ 
Fall, NY & 1.1 & 1.8\\ 
Spring, NY &1.3 &2.1 \\ 
\end{tabular}
\end{table}
\vspace{1.25cm}
\end{column}
\end{columns}

@@

%\textit{relate back to article shown in the beginning, which also mentions IMF forecast of 0.1 percent $\rightarrow$ almost 50\% of any predictive intervals at this time would actually be below zero}

\addtocounter{framenumber}{-1}
\begin{frame}{Making it Public - GitHub}
%Width of Prediction Intervals, by Horizon
%\textit{include graphic or table showing increasing uncertainty with rising horizons}\\ 
    \begin{figure}
        \centering
        \includegraphics[width=0.5\textwidth]{figures_svr/GitHub_Capture2.jpg}
        %\caption{Coverage at the 50 and 80 percent level}
        \label{fig:enter-label}
    \end{figure}
\end{frame}

\addtocounter{framenumber}{-1}
\begin{frame}{Making it Public - GitHub}
%Width of Prediction Intervals, by Horizon
%\textit{include graphic or table showing increasing uncertainty with rising horizons}\\ 
\begin{figure}
        \centering
        \includegraphics[width=0.6\textwidth]{figures_svr/GitHub_Capture1.jpg}
        %\caption{Coverage at the 50 and 80 percent level}
        \label{fig:enter-label}
    \end{figure}

\appendix
\beginbackup
\begin{frame}{WIS}
\begin{equation}
WIS_{\tau_{0:K}}(F,y) = \frac{1}{K + 1/2}\left(w_{0}|y-m| + \sum_{k=1}^{K}\left(w_k IS_{\tau_{k}}(F, y)\right)\right), 
\vspace*{5.5mm}
\end{equation}
\hspace*{3cm} with $w_k = (1-\tau)/2$
\end{frame}
\begin{frame}{Directional Errors - Scores}
\label{errorextraction}
\begin{figure}
        \centering
        \includegraphics[width=0.9\textwidth]{figures/wis_cpigdp_directional.pdf}
       %\caption{Coverage at the 50 and 80 percent level}
    \end{figure}
\end{frame}
\begin{frame}{Directional Errors - Coverage}
\begin{figure}
        \centering
        \includegraphics[width=0.9\textwidth]{figures/coverage_directional.pdf}
        %\caption{Coverage at the 50 and 80 percent level}
    \end{figure}
\end{frame}

@@

\end{document}
 
